---
title: "Median Age Vs. Healthcare Expense  Across Countries 2019 Comparative Analysis "
author: "Grant Gealy"
output:
  github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
health expenditure  from world bank
median age from kaggle via UN 
gdp per capita data from kaggle
 This project will focus on the relationship between the age of a society and other economic factors in order to better understand the economics of specifically how income level and average age influence factors such as gdp growth and government spending. Note (age data is from 2020 due to lack of data for every year )
```{r}
#load library 
#install.packages("reshape2")
library(tidyverse)
library(reshape2)
library(colorspace)


```
 
```{r}
#load data 
age= read.csv("data/MedianAge.csv")%>% melt()%>%rename(year = variable )%>%rename(age =value)%>%rename(country = Country)
#get rid of x in the year col
age$year= as.numeric(substr(age$year,2,5))
# pick to analyze 
nage=age%>%filter(year==2020) 
head(age)
```
```{r}
##load economic freedom data 
fre = read.csv("data/economic_freedom_index2019_data.csv")%>%rename(country = Country.Name)
fre$WEBNAME= NULL
fre
#take out names to make all data numeric then put the names back in 
numfre=sapply(fre[,-c(2,3)], function(x) as.numeric(gsub(
        "$","",x,fixed=TRUE))) 
frename=cbind(fre[,2:3],numfre) 
#check data type for every col
sapply(fre,class) 
head(fre)
```
```{r}
##load health data 
health = read.csv("data/health.csv")%>%select(-c(Country.Code,Indicator.Name, Indicator.Code))%>%rename(country =ï..Country.Name)%>%melt(id.vars= "country")%>%rename(year = variable)%>%rename(hexpenspergdp =value)%>%na.omit()
#get rid of x in the year col 
health$year= as.numeric(substr(health$year,2,5)) 
head(health)
#get the values just for 2020  
healthyear=health[health$year==2019,]

#health$var = as.numeric(gsub("([0-9]+).*$", "\\1", health$var))
#health =sapply(health,function(x) as.numeric(gsub("([0-9]+).*$", "\\1", x)))

```
```{r}
#make one data set From the separate data sets 

data=merge(frename,healthyear,"country",all.x = TRUE)
fulldata=merge(data,nage,"country",all.x = TRUE)%>%drop_na(c(age,hexpenspergdp))
```
start analysis exploratory
```{r}
#we found the summary statistics for age an, health expenses 
#and gdp growth rate 
fulldata%>%select(age,hexpenspergdp,GDP.Growth.Rate....
                  )%>%summary() 
#make histograms for variable of interest 
sapply(fulldata%>%select(age,hexpenspergdp,GDP.Growth.Rate....
                  ),hist) 
#find cor of variables of interest save it to a variable so we 
#can graph it 
test=fulldata%>%select(age,hexpenspergdp,GDP.Growth.Rate.... )%>%na.omit()%>%cor()%>%as.data.frame()
test=test%>%mutate(variable =row.names(test))%>%pivot_longer(cols = 0:3,names_to= "var2",values_to = "cor")

```
```{r}
selected=fulldata[fulldata$country=="Japan",]
#make a graph plotting age and health expenses/gdp to illustrate the relationship between the age of a society and the amount they investorsinvest in Healthcare 
ggplot(fulldata,aes(age,hexpenspergdp))+geom_point()+geom_smooth(method = lm)+geom_text(data = selected,mapping=aes(x=age,y=hexpenspergdp,label=country ),color="red")

```


```{r}


#make a graph plotting government integrity and health expenses I did this to show how corruption in poorer countries might exaggerate the health expenses actually going towards the people
ggplot(fulldata,aes(fulldata$Government.Integrity,hexpenspergdp))+geom_point(aes(size=GDP..Billions..PPP.,color=Region))+geom_smooth(method=lm)+labs(title = "health care spending vs government integrity 2019",y="Healthcare spending",x="government integrity ",size= "gdp in billions" ,color= "Region")+theme_bw()#+geom_text(data = selected,mapping=aes(x=Government.Integrity,y=hexpenspergdp,label=country ))



  #geom_text(label= fulldata$country)
```


```{r}
#make a graph showing the cor 
ggplot(test,aes(y=var2,x= variable, fill =cor))+geom_tile()+geom_text(aes(label= round(cor,4)), color ="blue ")+labs(title ="cor title ",y="y",x="x")+theme(legend.position = "none")+scale_fill_continuous_sequential(palette="red")
```
```{r}
ggplot(fulldata,aes(age,Business.Freedom))+geom_point()+geom_smooth(method="lm")+labs(title = "age vs. Business freedom",x="age",y=" Business freedom ")
```

```{r}

agegdp=lm(fulldata$hexpenspergdp~age,data = fulldata)
summary(agegdp)
agegdp$coefficients
```


```{r}
usa=which(fulldata$country=="United States")
agegdp$residuals[usa]

```


```{r}
#as a society ages they spend more on health 

#can we also do the growth rate 
#as yes that's what I am thinking 

```
```{r}

```



 
 
 
 
 |

